---
title: "Útrokningar av hvussu nógvir bitlar verða upptiknir á kraklingaaliøkjunum"
author: "Birgitta Andreasen"
date: "22. jul 2021"
output: 
  bookdown::html_document2:
    toc: true
    toc_float: true
    number_sections: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)

# pakkar ið skulu nýtast
library(tidyverse)
library(knitr)
library(kableExtra)
# Suppress summarise info
options(dplyr.summarise.inform = FALSE)

# scripts sum skulu koyrast
source("scripts/func_kraklingur_ingested_incl_sensitivity_analysis.R")

# skal tú gera nýggjar df?
nyggjar <- TRUE

# um tú ikki hevur nokk av ram, so ber ikki til at hyggja at rádata plottunum, set tí rawdata <- FALSE
rawdata <- TRUE


#plot labels
frac_assimilated_label <- "Waste assimilated"
frac_ingested_label <- "Particles ingested"
speedfactor_label <- "Factor adjusted current speed"
volfilt <- "FR[v]~(l~m^{-3}~h^{-1}))"
FRv_label <- expression(FR[italic(v)]~(italic(l~m^{-3}~h^{-1})))
FRv_label_long <- expression("Volumetric Filtration Rate, "(italic(l~m^{-3}~h^{-1})))
Sv_label <- expression("Settling Velocity, "(italic(cm~s^{-1})))
Sv_label_standard <- expression(italic(v)~(italic(m~s^{-1})))

# stødd í mm á myndunum
singlepanel <- 81
doublepanel <- 169

```

# Útrokningar

## Upplýsingar

Four settling velocities were modeled 7.5, 3.2, 1.5, and 0.1 cm s^-1^ at a distribution of 65, 20, 10, and 5 % respectively.

The ingestion of fish waste (particle waste (PW)) by the blue mussels $I_{PW}$ can be expressed as Equation \@ref(eq:ingestion)

```{=tex}
\begin{equation}
I_{PW} = FR_{v} C (\#eq:ingestion)
\end{equation}
```
where:

-   $FR_{v}$ is the filtration rate of the entire mussle population per $m^{3}$ mussel farm
-   $C$ is the concentration of particles

The change in mass of particles ($m_{PW}$) due to ingestion by the mussels during the total time the waste particles are in the mussel farm ($t$), can be expressed as Equation \@ref(eq:changemass).

```{=tex}
\begin{equation}
\frac{\partial m_{PW}}{\partial t} = -FR_{v} m_{PW} (\#eq:changemass)
\end{equation}
```
```{=tex}
\begin{equation}
m_{PW}(t) = m_{PW}(0) e^{-FR_{v} t} (\#eq:massnull)
\end{equation}
```
where $m_{PW}(0)$ is the mass of particles at time 0.

When taking into account the constant sinking velocity ($S_{v}$) of the particles, the mass of particles can be expressed as a function of the vertical distance the particles have covered in the mussel farm ($z$):

```{=tex}
\begin{equation}
m_{PW}(z) = m_{PW}(0) e^{-\frac{FR_{v} z}{S_{v}}} (\#eq:verticaldistz)
\end{equation}
```
In the dispersion model, the horizontal position of the particle is logged per 1 cm. Thereby the final equation for the fraction of fish waste that is assimilated by the blue mussels during passage through the mussel farm can be expressed as:

```{=tex}
\begin{equation}
\frac{m_{assimilated}}{PW} = AE \displaystyle\sum_{S_{v}}^{} \left( \frac{\alpha_{S_v}}{\sum_{i = 0}^{M} f(i, S_{v})} \displaystyle\sum_{i=0}^{M} \left[ f(i, S_{v}) \left( 1 - e^{-\frac{FR_{v}}{S_{v}} z(i)} \right) \right] \right) (\#eq:assimilated)
\end{equation}
```
where:

-   i describes the amount of vertical positions in the mussel farm where the particle was observed
-   $f(i,S_v)$ describes the number of particle with sinking speed $S_v$ observed on $i$ positions
-   M is the maximum possible number that $i$ can obtain
-   $\alpha S_v$ is the fraction of PW with the sinking velocity $S_v$
-   $z(i)$ is the vertical distance covered in the farm given it was observed $i$ times

Equation \@ref(eq:assimilated) written differently to match dataframe columns.

```{=tex}
\begin{equation}
\frac{m_{assimilated}}{PW} = AE \displaystyle\sum_{S_{v}}^{} \left( \alpha_{S_v} \frac{\displaystyle\sum_{i=0}^{M} \left[ f(i, S_{v}) \left( 1 - e^{-\frac{FR_{v}}{S_{v}} z(i)} \right) \right]}{\sum_{i = 0}^{M} f(i, S_{v})} \right) (\#eq:assimilated1)
\end{equation}
```
```{=tex}
\begin{equation}
frac\_assimilated = AE \displaystyle\sum_{S_{v}}^{} \left( alpha\_Sv \frac{N\_ingested\_Sv}{N\_dropped} \right) (\#eq:assimilated2)
\end{equation}
```


```{r}
# hvør/hvørjar søkkiferðir skal koyrast, vel "01", "15", "32" ella "75" (string)
ferdir <- c("01", "15", "32", "75")

# hvussu skulu ferðirnar býtast / alpha
byti <- data.frame(byti1 = c(0, 0.1, 0.2, 0.7), 
                   byti2 = c(0.05, 0.1, 0.2, 0.65),
                   byti3 = c(0.27, 0.1, 0.2, 0.43))
bytidf <- byti %>% 
  cbind(ferdir) %>% 
  pivot_longer(-ferdir) %>% 
  rename(ferd = ferdir,
         bytinavn = name,
         alpha_Sv = value)

# hvussu eita kraklingaaliøkini í fílunum?
krak_navn <- c("1", "2")

# hvat label ynskir tú at seta á kraklingaaliøkini, skal hava somu raðfylgi sum krak_navn omanfyri!!
krak_label <- c("surface", "submerged")

df_kraklabel <- data.frame(kraknavn = krak_navn, 
                           label = factor(krak_label, levels = krak_label))

# hvussu nógvir bitlar eru í hvørjari slepping?
particles <- 58313

# hvussu nógv drops vóru?
drops <- 1028

totalparticles <- drops * particles

# to convert L/h to m3/s, have to divide by (1000*3600)
lhm3s <- 3600000

# clearingsrate (L/(h*individual))
CR <- 3.2

# partur av tilfari sum kræklingur far upptikið
AE <- 0.86

# mussel density [mussel/m]
dm <- c(117.1875, seq(120, 2500, 10))

# distance between droppers [m]
dropperdist <- 0.5

# filtration rate [L/(m*h)]
FR = CR * dm

# volumetric filtration rate FRv [L/(m3 h)] -> [m3/(m3*s)]
FRv <- CR * dm * 1/dropperdist / lhm3s

#FRv * lhm3s



```

## Samla dáturnar fyri kraklingaaliøkini við allar ferðir

```{r}

speedfactor <- sprintf("%.2f", seq(0.1,4,0.1))

if(nyggjar) {
  
  # !obs you have to gather them in a single folder! or rewrite previous code to do this for you!!
  summaries_folder <- "data/processed/summaries/"
  
  move_summaries(ferdir, krak_navn, speedfactor, summaries_folder)
  
  for(i in speedfactor) {
    
    kraksamla(folderpath = summaries_folder, ferdir =  ferdir, krak_navn = krak_navn, speedfactor = i)
  }
  
}


```

## hyggja at rá data (speedfactor = 1.00)

```{r}
# innles nýggju daturammuna, dataramman bleiv goymd her:
kraksamla1 <- read_csv("data/processed/sensitivities/krak_particle_summary_1.00.csv", 
                       col_types = cols(kraknavn = col_character()), lazy = FALSE)

# datawrangling
kraksamla1 <- kraksamla1 %>% 
  left_join(data.frame(kraknavn = krak_navn,
                       label = krak_label), 
            by = "kraknavn") %>% 
  mutate(meters = mm/1000,
         speed = as.numeric(ferd)/1000)


# summary
summaryparticles <- kraksamla1 %>% 
  group_by(kraknavn, label, speed, ferd) %>% 
  summarise(bitlar_aldrin_alibruk = sum(count[mm==0]),
            bitlar_igjognum_alibruk = sum(count[mm>0]),
            bitlar_total = sum(count),
            frac_igjognum = bitlar_igjognum_alibruk/bitlar_total) %>% 
  ungroup()

byti_summary <- summaryparticles %>% 
  left_join(bytidf) %>% 
  group_by(label, bytinavn) %>% 
  summarise(N_dropped = sum(alpha_Sv*bitlar_total),
            N_reach_musselfarm = sum(alpha_Sv*bitlar_igjognum_alibruk),
            perc_reach_musselfarm = N_reach_musselfarm/N_dropped*100)

if(rawdata == FALSE){
  rm(kraksamla1)
}
```

### Kraklinga aliøki undir liðini á laksaalingini

**legg til merkis at y-axin er ikki á sama scala!**

```{r}

oki <- "1"

plot1 <- kraksamla1 %>% 
  filter(mm != 0) %>% 
  filter(mm < 10000) %>% 
  filter(kraknavn == oki) %>% 
  ggplot(aes(meters, count)) +
  geom_point() +
  facet_grid(speed~kraknavn+label, scales = "free_y") +
  geom_text(data = summaryparticles %>% 
              filter(kraknavn == oki),
            aes(x = Inf, y = Inf, label = paste0("particles that go through the mussle farm = ", scales::percent(frac_igjognum, 0.01))), 
            hjust = "inward", vjust = "inward", size = 3, colour = "darkred")

#ggsave(paste0("images/particles_pr_length_krak_", oki, "_1.00.png"), plot1, width = 6, height = 8)

plot1

```

### Kraklinga aliøki undir laksaalingini

**legg til merkis at y-axin er ikki á sama scala! og at bitlar, ið detta beint igjognum kraklingaaliøki ikki eru vístir!!**

```{r eval}

oki <- "2"

plot1 <- kraksamla1 %>% 
  filter(mm != 0) %>% 
  filter(mm <= 10000) %>% 
  filter(kraknavn == oki) %>% 
  ggplot(aes(meters, count)) +
  geom_point() +
  facet_grid(speed~kraknavn+label, scales = "free_y") +
  geom_text(data = summaryparticles %>% 
              filter(kraknavn == oki),
            aes(x = Inf, y = Inf, label = paste0("particles that go through the mussle farm = ", scales::percent(frac_igjognum, 0.01))), 
            hjust = "inward", vjust = "inward", size = 3, colour = "darkred")

#ggsave(paste0("images/particles_pr_length_krak_", oki, "_1.00.png"), plot1, width = 6, height = 8)

plot1
```

```{r}
kbl(summaryparticles) %>% 
  kable_styling() %>% 
  row_spec(1:4, background = "#42f5b9") %>% 
  row_spec(5:8, background = "#f783bb")

```

# PW ingested

```{r}
# ingested by mussels
speedfactor <- c("0.00", speedfactor)

sensitivity_at_zero(ferdir, krak_navn, totalparticles)

ingested <- krakingested(ferdir, speedfactor, FRv)

```

```{r}

# vel ein FRv
filterFRv <- CR * 600 * 1/dropperdist / lhm3s

summaryingested <- ingested %>% 
  filter(speedfactor == 1.00) %>% 
  filter(FRv == filterFRv)

kbl(summaryingested) %>% 
  kable_styling() %>% 
  row_spec(1:4, background = "#42f5b9") %>% 
  row_spec(5:8, background = "#f783bb")
```

## plot frac_ingested vs. speedfactor at one FRv

```{r}

ingestedspeedfactor <- ingested %>% 
  left_join(df_kraklabel, by = "kraknavn") %>%  
  filter(FRv == filterFRv) %>% 
  ggplot(aes(speedfactor, frac_ingested, colour = factor(Sv))) +
  geom_hline(yintercept = 0, colour = "grey20")+
  geom_line(size = 0.6, aes(linetype = factor(Sv))) +
  scale_color_viridis_d(name = Sv_label_standard) +
  scale_linetype_discrete(name = Sv_label_standard) +
  facet_wrap(.~label, scales = "free_y", ncol = 1) +
  xlab(speedfactor_label) +
  ylab(frac_ingested_label) +
  guides(colour = guide_legend(nrow = 1, byrow = TRUE, title.position = "top")) +
  #geom_vline(xintercept = filterFRv*lhm3s, linetype = 2) +
  #geom_text(data = data.frame(filterFRv), 
  #          aes(x = filterFRv*lhm3s, y = Inf,
  #              label = paste0(filterFRv*lhm3s, "~L~m^-3~h^-1")),
  #          vjust = "inward", hjust = 0, colour = "black", parse = TRUE) +
  theme_minimal(base_size = 9) +
  theme(strip.background = element_rect(fill = "grey90", colour = NA),
        legend.position = "top")

ingestedspeedfactor

ingestedspeedfactor2 <- ingestedspeedfactor +
  facet_wrap(.~label, scales = "free_y", ncol = 2) +
  guides(colour = guide_legend(ncol = 1)) +
  theme(legend.position = "right")


ggsave(plot = ingestedspeedfactor, 
       filename = "images/particle_ingestion_vs_speedfactor_at_Sv_singlepanel.pdf",
       units = "mm",
       width = singlepanel, height = 160)

ggsave(plot = ingestedspeedfactor2, 
       filename = "images/particle_ingestion_vs_speedfactor_at_Sv_doublepanel.pdf",
       units = "mm",
       width = doublepanel, height = 80)

```

## plot frac_ingested vs. FRv at speedfactor 1.00

```{r}

filterspeedfactor <- 1.00

particleingestion <- ingested %>% 
  filter(speedfactor == filterspeedfactor) %>% 
  mutate(frveind = FRv*lhm3s) %>% 
  left_join(df_kraklabel, by = "kraknavn") %>% 
  ggplot(aes(frveind, frac_ingested, colour = factor(Sv))) +
  geom_hline(yintercept = 0, colour = "grey20")+
  geom_line(size = 1, aes(linetype = factor(Sv))) +
  scale_color_viridis_d(name = Sv_label_standard) +
  scale_linetype_discrete(name = Sv_label_standard) +
  facet_wrap(.~label, scales = "free_y", ncol = 1) +
  xlab(FRv_label_long) +
  ylab(frac_ingested_label) +
  guides(colour = guide_legend(nrow = 2, byrow = TRUE)) +
  #geom_vline(xintercept = filterFRv*lhm3s, linetype = 2) +
  #geom_text(data = data.frame(filterFRv), 
  #          aes(x = filterFRv*lhm3s, y = Inf,
  #              label = paste0(filterFRv*lhm3s, "~L~m^-3~h^-1")),
  #          vjust = "inward", hjust = 0, colour = "black", parse = TRUE) +
  theme_minimal(base_size = 9) +
  theme(strip.background = element_rect(fill = "grey90", colour = NA),
        legend.position = "bottom")

particleingestion2 <- particleingestion +
  facet_wrap(.~label, scales = "free_y", ncol = 2) +
  guides(colour = guide_legend(ncol = 1)) +
  theme(legend.position = "right")

# ggsave(plot = particleingestion, 
#        filename = "images/particle_ingestion_vs_FRv_at_Sv_singlepanel.pdf",
#        units = "mm",
#        width = singlepanel, height = 160)
# 
# ggsave(plot = particleingestion2, 
#        filename = "images/particle_ingestion_vs_FRv_at_Sv_doublepanel.pdf",
#        units = "mm",
#        width = doublepanel, height = 80)

```

## fraktion assimilated

```{r}

assimilated <- krakassimilated(bytidf, AE, ingested)

head(assimilated)
```

## Maximum assimilated

```{r}
# vel ein FRv
filterFRv <- CR * 600 * 1/dropperdist / lhm3s

maxsummary <- assimilated %>% 
  group_by(kraknavn, bytinavn) %>% 
  filter(frac_assimilated == max(frac_assimilated) | (speedfactor %in% c(0.8,1) & FRv == filterFRv)) %>% 
  left_join(df_kraklabel) %>% 
  ungroup() %>% 
  mutate(perc_reach_musselfarm = N_kraklingaoki/N_dropped*100,
         perc_assimilated = frac_assimilated*100) %>% 
  select(label, bytinavn, byti, ferdir, speedfactor, FRv, AE, perc_reach_musselfarm, perc_assimilated) %>% 
  arrange(FRv,speedfactor, label, bytinavn)

maxsummary
```

## plot frac_assimilated vs. speedfactor at one FRv

```{r}

assimilated_pr_ferd <- krakassimilated_individual(AE, ingested)

litir <- c("#440154FF", "#31688EFF", "#35B779FF", "#fcbe03FF")


assimilatedspeedfactor <- assimilated_pr_ferd %>% 
  left_join(df_kraklabel, by = "kraknavn") %>%  
  filter(FRv == filterFRv) %>%  
  ggplot(aes(speedfactor, frac_assimilated, colour = factor(Sv*100))) +
  geom_hline(yintercept = 0, colour = "grey60")+
  geom_line(size = 0.6, aes(linetype = factor(Sv*100))) +
  #scale_color_viridis_d(name = Sv_label, end = 0.9, direction = 1) +
  scale_color_manual(values = litir, name = Sv_label) +
  scale_linetype_discrete(name = Sv_label) +
  facet_wrap(.~label, scales = "free_y", ncol = 1) +
  xlab(speedfactor_label) +
  ylab(frac_assimilated_label) +
  scale_y_continuous(labels = scales::percent) +
  guides(colour = guide_legend(nrow = 1, byrow = TRUE, title.position = "top")) +
  #geom_vline(xintercept = filterFRv*lhm3s, linetype = 2) +
  #geom_text(data = data.frame(filterFRv), 
  #          aes(x = filterFRv*lhm3s, y = Inf,
  #              label = paste0(filterFRv*lhm3s, "~L~m^-3~h^-1")),
  #          vjust = "inward", hjust = 0, colour = "black", parse = TRUE) +
  theme_minimal(base_size = 9) +
  theme(strip.background = element_rect(fill = "grey90", colour = NA),
        legend.position = "top")

assimilatedspeedfactor

assimilatedspeedfactor2 <- assimilatedspeedfactor +
  facet_wrap(.~label, scales = "free_y", ncol = 2) +
  #guides(colour = guide_legend(ncol = 1)) +
  theme(legend.position = "top")


ggsave(plot = assimilatedspeedfactor, 
       filename = "images/waste_assimilatd_vs_speedfactor_at_Sv_singlepanel.pdf",
       units = "mm",
       width = singlepanel, height = 160)

ggsave(plot = assimilatedspeedfactor2, 
       filename = "images/waste_assimilated_vs_speedfactor_at_Sv_doublepanel.pdf",
       units = "mm",
       width = doublepanel, height = 80)

ggsave(plot = assimilatedspeedfactor, 
       filename = "images/waste_assimilatd_vs_speedfactor_at_Sv_singlepanel.png",
       units = "mm",
       width = singlepanel, height = 160)

ggsave(plot = assimilatedspeedfactor2, 
       filename = "images/waste_assimilated_vs_speedfactor_at_Sv_doublepanel.png",
       units = "mm",
       width = doublepanel, height = 80)

```

## plot frac_assimilated vs. speedfactor at all calculated FRv and at three different speed divisions a, b, and c

```{r}

colorbreaks <- seq(0,17500, 2500)
colorbreaks <- c(750, seq(2500, 15000, 2500), 16000)
dflabels <- data.frame(pent = factor(colorbreaks),
                       bolkurgroup = factor(1:length(colorbreaks)))

ordid <- "division "

ordiorder <- c(paste0(krak_label[[1]], " - ", ordid, c("a", "b", "c")), paste0(
  krak_label[[2]], " - ", ordid, c("a", "b", "c")))

dfplot <- assimilated %>%
  left_join(df_kraklabel) %>% 
  ungroup() %>% 
  mutate(bolkur = FRv*lhm3s,
         bolkurgroup = cut(bolkur, colorbreaks, labels = FALSE),
         bolkurgroup = factor(bolkurgroup),
         disk = cut_width(bolkur, width = 4000),
         #bolkur = factor(bolkur),
         #bolkur = fct_reorder(bolkur, FRv, .desc = TRUE)
  ) %>% 
  mutate(combination = str_replace_all(bytinavn, "byti", ordid),
         combination = str_replace_all(combination, c("1" = "a",
                                                      "2" = "b",
                                                      "3" = "c")),
         wrapabel = paste0(label, " - ", combination),
         wrapabel = factor(wrapabel, levels = ordiorder)) %>% 
  left_join(dflabels) %>% 
  mutate(pentfactor = factor(pent, levels = rev(dflabels$pent)))

fill.order = factor(dfplot$pent, levels = colorbreaks)

filteras <- dfplot %>% 
  filter(FRv == filterFRv)
filterspeed <- dfplot %>% 
  filter(speedfactor == 1.00) %>% 
  filter(FRv <= filterFRv)

dfscales <- data.frame(kraknavn = as.character(c(1, 1, 1, 2, 2, 2)),
                       bytinavn =c("byti1", "byti2", "byti3", "byti1", "byti2", "byti3"),
                       x = c(3,3,3,3,3,3),
                       y = c(0.024, 0.024, 0.065, 0.65, 0.65, 0.65)) %>% 
  left_join(filterspeed %>% filter(FRv == filterFRv)) %>% 
  left_join(df_kraklabel) %>% 
  mutate(combination = str_replace_all(bytinavn, "byti", ordid),
         combination = str_replace_all(combination, c("1" = "a",
                                                      "2" = "b",
                                                      "3" = "c")),
         wrapabel = paste0(label, " - ", combination),
         wrapabel = factor(wrapabel, levels = ordiorder))



different_fractions <- dfplot %>% 
  ggplot(aes(y = frac_assimilated, x = speedfactor, 
             color = bolkur, 
             fill = bolkur
  )) +
  #scale_color_stepsn(colors = viridis::viridis(8), name = FRv_label)+
  #scale_fill_stepsn(colors = viridis::viridis(8), name = FRv_label)+
  scale_color_binned(type = "viridis", n.breaks = 7, 
                     name = FRv_label_long, 
                     #breaks = colorbreaks[1:7],
                     limits = c(750, 16000)) +
  #scale_color_viridis_d(name = FRv_label)+
  #scale_fill_viridis_d(name = FRv_label, ) +
  
  #geom_area(aes(group= (FRv)), position = position_identity()) +
  geom_line(aes(group = FRv), size = 0.45) +
  geom_line(data = filteras, colour = "red", show.legend = FALSE) +
  #geom_line(data = filterspeed, colour = "red", show.legend = FALSE) +
  geom_point(data = filterspeed %>% filter(FRv == filterFRv), 
             colour = "red", show.legend = FALSE) +
  geom_blank(data = dfscales, aes(x,y))+
  ggrepel::geom_text_repel(data = dfscales, 
                           aes(label = scales::percent(frac_assimilated, 0.01)),
                           nudge_x = Inf, nudge_y = Inf, 
                           segment.linetype = 3,
                           size = 3,
                           #segment.curvature = -0.001,
                           #arrow = arrow(length = unit(0.1, "cm")),
                           colour = "red"
  ) +
  facet_wrap(wrapabel~., scales = "free")+
  guides(color = guide_colorbar(barheight = 0.5, barwidth = 15, 
                                title.position = "top"))+
  ylab(frac_assimilated_label)+
  scale_y_continuous(labels = scales::percent, expand = c(0,0)) +
  #coord_cartesian(xlim = c(0.11,4), expand = FALSE)+
  xlab(speedfactor_label) +
  geom_hline(yintercept = 0, colour = "grey60") +
  theme_minimal(base_size = 9) +
  theme(legend.position = "top",
        legend.title = element_text(),
        strip.background = element_rect(fill = "grey90", colour = NA))

different_fractions


```

```{r}
ggsave(plot = different_fractions, 
       filename = "images/waste_assimilated_vs_speedfactor_at_FRv_doublepanel.pdf",
       units = "mm",
       width = doublepanel, height = 120)

ggsave(plot = different_fractions, 
       filename = "images/waste_assimilated_vs_speedfactor_at_FRv_doublepanel.png",
       units = "mm",
       width = doublepanel, height = 120)


```
